---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
devtools::load_all(".", quiet = TRUE)
```

# `rw`: Robins-Wang variance estimation for multiple imputation

<!-- badges: start -->
[![R-CMD-check](https://github.com/yu000434/RW-package-test/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/yu000434/RW-package-test/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `rw` package computes [Robins-Wang variance
estimates](https://doi.org/10.1093/biomet/87.1.113) for analyses of multiply
imputed data. It supports parametric chained-equation imputation and predictive
mean matching with donor-source correction.

## Installation

The package requires the development version of `mice` that records fitted
imputation models through `tasks = "train"`.

```r
remotes::install_github("amices/mice@c9b67ae1cd54784267a01f1b70d93a40d509a5de")
remotes::install_github("yu000434/RW-package-test")
```

## Parametric imputation

This example imputes the incomplete variables in the `nhanes` data using
normal working models.

```{r parametric, message=FALSE}
library(rw)
library(mice)

nhanes_scaled <- as.data.frame(scale(nhanes))
method <- make.method(nhanes_scaled)
method[] <- ""
method[c("bmi", "hyp", "chl")] <- "norm"

set.seed(1)
imp_norm <- mice(nhanes_scaled, method = method, m = 5,
                 tasks = "train", print = FALSE)
fit_norm <- with_rw(imp_norm, lm(bmi ~ age + hyp))
pool_rw(fit_norm)
```

For comparison, Rubin's rules can be applied to the same imputations.

```{r rubin-parametric}
pool(with(imp_norm, lm(bmi ~ age + hyp))) |>
  summary()
```

## Predictive mean matching

The `pmmrw` method uses the ordinary MICE donor draw and records the selected
donor. The completed values are unchanged. The matching derivative is used
only for variance estimation, and `pool_rw()` accounts for repeated use of the
same donor.

Use `tasks = "train"` with numeric PMM variables and `matchtype = 1`.
The options `exclude = NULL`, `use.matcher = FALSE`, and `mlocal = 1` remain
at their defaults. The donor pool size is controlled by `donors` (default 5).

```{r pmm, message=FALSE}
nhanes_pmm <- nhanes_scaled[c("age", "bmi")]
method <- make.method(nhanes_pmm)
method[] <- ""
method["bmi"] <- "pmmrw"

set.seed(2)
imp_pmm <- mice(nhanes_pmm, method = method, m = 5,
                tasks = "train", print = FALSE)
fit_pmm <- with_rw(imp_pmm, lm(bmi ~ age))
pool_rw(fit_pmm)
```

Using the same seed with ordinary MICE PMM gives the same completed values.

```{r pmm-parity, message=FALSE}
method["bmi"] <- "pmm"
set.seed(2)
imp_standard <- mice(nhanes_pmm, method = method, m = 5, print = FALSE)
identical(imp_pmm$imp$bmi, imp_standard$imp$bmi)
```

The donor IDs used in each imputation are available when needed.

```{r donors}
head(extract_donor_id(imp_pmm, "bmi"))
```

## Giganti and Shepherd example

The package includes the simulated `giganti_data`. The first 500 observations
are used here to keep the example short. Both `A` and `D` are incomplete.

```{r gs-setup}
data("giganti_data", package = "rw")
gs <- giganti_data[seq_len(500), ]
predictors <- c("X1", "X2", "A.star", "D.star")

pred <- make.predictorMatrix(gs)
pred[,] <- 0L
pred["A", predictors] <- 1L
pred["D", c(predictors, "A")] <- 1L
```

With parametric imputation, `A` uses a normal model and `D` uses logistic
regression.

```{r gs-parametric, message=FALSE}
method <- make.method(gs)
method[] <- ""
method[c("A", "D")] <- c("norm", "logreg")

set.seed(41)
imp_gs_norm <- mice(gs, m = 2, method = method, predictorMatrix = pred,
                    tasks = "train", print = FALSE)
fit_gs_norm <- with_rw(
  imp_gs_norm,
  glm(D ~ A, subset = A > 2, family = binomial())
)
pool_rw(fit_gs_norm)
```

For predictive mean matching, `pmm_kappa_binomial()` computes the PMM cross
term for the binomial analysis before the final variance is assembled.

```{r gs-pmm, message=FALSE}
method[c("A", "D")] <- c("pmmrw", "logreg")

set.seed(42)
imp_gs_pmm <- mice(gs, m = 2, method = method, predictorMatrix = pred,
                   tasks = "train", print = FALSE)
fit_gs_pmm <- with_rw(
  imp_gs_pmm,
  glm(D ~ A, subset = A > 2, family = binomial())
)
kappa <- pmm_kappa_binomial(fit_gs_pmm, "A", threshold = 2)
pool_rw(fit_gs_pmm, pmm_kappa = kappa)
```

## Current scope

The package supports `norm` and `logreg` imputation with unweighted `lm()` or
binomial `glm()` analysis using the default logit link.

PMM variance estimation currently requires one numeric `pmmrw` variable with
fully observed predictors in its matching model:

- `pool_rw()` automatically handles an untransformed PMM response in `lm()`.
  Analysis predictors and any subset must depend only on fully observed variables.
- `pmm_kappa_binomial()` handles `glm(D ~ A, family = binomial())`, where `A`
  uses PMM and `D` uses `logreg`. The analysis can use all rows or the subset
  `A > threshold`. The `logreg` model must include `A` as an untransformed
  additive predictor; its other predictors must be fully observed.
- Other PMM analyses need an expected-score function supplied to `pmm_kappa()`.
  Its arguments and required output are described in `?pmm_kappa`.

## Source files

| File | Role |
|:-----|:-----|
| `R/pmmrw.R` | Standard MICE PMM draws and donor recording. |
| `R/parametric.R` | Imputation scores and influence contributions for `norm` and `logreg`. |
| `R/pmm.R` | PMM matching probabilities, derivatives, and expected-score cross terms. |
| `R/analysis.R` | Analysis-model scores and their derivatives. |
| `R/rw.R` | Fitting, RW variance assembly, pooling, and result methods. |

Function comments in `R/` are kept short; the full reference pages are maintained
in `man/`. `devtools::document()` updates `NAMESPACE` only. Edit `README.Rmd`
and run `rmarkdown::render("README.Rmd")` to update this page and its example
output. The package website is built from this README and the reference pages.
