---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
devtools::load_all(".", quiet = TRUE)
```

# `rw`: Robins-Wang variance estimation for multiple imputation

The `rw` package computes [Robins-Wang variance
estimates](https://doi.org/10.1093/biomet/87.1.113) for analyses of multiply
imputed data. It supports parametric chained-equation imputation and predictive
mean matching with donor-source correction.

## Installation

The package requires the development version of `mice` that records fitted
imputation models through `tasks = "train"`.

```r
remotes::install_github("amices/mice@c9b67ae1cd54784267a01f1b70d93a40d509a5de")
remotes::install_git("git@github.com:yu000434/RW-package-test.git")
```

## Parametric imputation

This example imputes the incomplete variables in the `nhanes` data using
normal working models.

```{r parametric, message=FALSE}
library(rw)
library(mice)

nhanes_scaled <- as.data.frame(scale(nhanes))
method <- make.method(nhanes_scaled)
method[] <- ""
method[c("bmi", "hyp", "chl")] <- "norm"

set.seed(1)
imp_norm <- mice(nhanes_scaled, method = method, m = 5,
                 tasks = "train", print = FALSE)
fit_norm <- with_rw(imp_norm, lm(bmi ~ age + hyp))
pool_rw(fit_norm)
```

For comparison, Rubin's rules can be applied to the same imputations.

```{r rubin-parametric}
pool(with(imp_norm, lm(bmi ~ age + hyp))) |>
  summary()
```

## Predictive mean matching

The `pmmrw` method uses the ordinary MICE donor draw and records the selected
donor. The completed values are unchanged. The matching derivative is used
only for variance estimation, and `pool_rw()` accounts for repeated use of the
same donor.

```{r pmm, message=FALSE}
nhanes_pmm <- nhanes_scaled[c("age", "bmi")]
method <- make.method(nhanes_pmm)
method[] <- ""
method["bmi"] <- "pmmrw"

set.seed(2)
imp_pmm <- mice(nhanes_pmm, method = method, m = 5,
                tasks = "train", print = FALSE)
fit_pmm <- with_rw(imp_pmm, lm(bmi ~ age))
pool_rw(fit_pmm)
```

Using the same seed with ordinary MICE PMM gives the same completed values.

```{r pmm-parity, message=FALSE}
method["bmi"] <- "pmm"
set.seed(2)
imp_standard <- mice(nhanes_pmm, method = method, m = 5, print = FALSE)
identical(imp_pmm$imp$bmi, imp_standard$imp$bmi)
```

The donor IDs used in each imputation are available when needed.

```{r donors}
head(extract_donor_id(imp_pmm, "bmi"))
```

## Giganti and Shepherd example

The package includes the simulated `giganti_data`. The first 500 observations
are used here to keep the example short. Both `A` and `D` are incomplete.

```{r gs-setup}
data("giganti_data", package = "rw")
gs <- giganti_data[seq_len(500), ]
predictors <- c("X1", "X2", "A.star", "D.star")

pred <- make.predictorMatrix(gs)
pred[,] <- 0L
pred["A", predictors] <- 1L
pred["D", c(predictors, "A")] <- 1L
```

With parametric imputation, `A` uses a normal model and `D` uses logistic
regression.

```{r gs-parametric, message=FALSE}
method <- make.method(gs)
method[] <- ""
method[c("A", "D")] <- c("norm", "logreg")

set.seed(41)
imp_gs_norm <- mice(gs, m = 2, method = method, predictorMatrix = pred,
                    tasks = "train", print = FALSE)
fit_gs_norm <- with_rw(
  imp_gs_norm,
  glm(D ~ A, subset = A > 2, family = binomial())
)
pool_rw(fit_gs_norm)
```

For predictive mean matching, `pmm_kappa_binomial()` computes the PMM cross
term for the binomial analysis before the final variance is assembled.

```{r gs-pmm, message=FALSE}
method[c("A", "D")] <- c("pmmrw", "logreg")

set.seed(42)
imp_gs_pmm <- mice(gs, m = 2, method = method, predictorMatrix = pred,
                   tasks = "train", print = FALSE)
fit_gs_pmm <- with_rw(
  imp_gs_pmm,
  glm(D ~ A, subset = A > 2, family = binomial())
)
kappa <- pmm_kappa_binomial(fit_gs_pmm, "A", threshold = 2)
pool_rw(fit_gs_pmm, pmm_kappa = kappa)
```

## Current scope

The package supports `norm` and `logreg` imputation with `lm()` or binomial
`glm()` analysis. PMM pooling supports one numeric `pmmrw` variable as the
response of `lm()` or as the predictor in the documented binomial analysis.
Other PMM analyses can define their expected analysis score through
`pmm_kappa()`.
