---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
devtools::load_all(".", quiet = TRUE)
```

# `rw`: Robins-Wang variance estimation for multiple imputation

The `rw` package computes [Robins-Wang variance
estimates](https://doi.org/10.1093/biomet/87.1.113) for analyses of multiply
imputed data. It supports parametric chained-equation imputation and predictive
mean matching with donor-source correction.

## Installation

The package requires the development version of `mice` that records fitted
imputation models through `tasks = "train"`.

```r
remotes::install_github("amices/mice@c9b67ae1cd54784267a01f1b70d93a40d509a5de")
remotes::install_git("git@github.com:yu000434/RW-package-test.git")
```

## Parametric imputation

This example imputes the incomplete variables in the `nhanes` data using
normal working models.

```{r parametric, message=FALSE}
library(rw)
library(mice)

nhanes_scaled <- as.data.frame(scale(nhanes))
method <- make.method(nhanes_scaled)
method[] <- ""
method[c("bmi", "hyp", "chl")] <- "norm"

set.seed(1)
imp_norm <- mice(nhanes_scaled, method = method, m = 5,
                 tasks = "train", print = FALSE)
fit_norm <- with_rw(imp_norm, lm(bmi ~ age + hyp))
pool_rw(fit_norm)
```

For comparison, Rubin's rules can be applied to the same imputations.

```{r rubin-parametric}
pool(with(imp_norm, lm(bmi ~ age + hyp))) |>
  summary()
```

## Predictive mean matching

The `pmmrw` method uses the ordinary MICE donor draw and records the selected
donor. The completed values are unchanged. The matching derivative is used
only for variance estimation, and `pool_rw()` accounts for repeated use of the
same donor.

```{r pmm, message=FALSE}
nhanes_pmm <- nhanes_scaled[c("age", "bmi")]
method <- make.method(nhanes_pmm)
method[] <- ""
method["bmi"] <- "pmmrw"

set.seed(2)
imp_pmm <- mice(nhanes_pmm, method = method, m = 5,
                tasks = "train", print = FALSE)
fit_pmm <- with_rw(imp_pmm, lm(bmi ~ age))
pool_rw(fit_pmm)
```

Using the same seed with ordinary MICE PMM gives the same completed values.

```{r pmm-parity, message=FALSE}
method["bmi"] <- "pmm"
set.seed(2)
imp_standard <- mice(nhanes_pmm, method = method, m = 5, print = FALSE)
identical(imp_pmm$imp$bmi, imp_standard$imp$bmi)
```

The donor IDs used in each imputation are available when needed.

```{r donors}
head(extract_donor_id(imp_pmm, "bmi"))
```

## Current scope

The package supports `norm` and `logreg` imputation with `lm()` or binomial
`glm()` analysis. Automatic PMM pooling currently supports one numeric
`pmmrw` variable used as the response of `lm()`. More complex PMM analyses can
provide their cross-term matrix through the `pmm_kappa` argument to
`pool_rw()`.
