---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
devtools::load_all(".", quiet = TRUE)
```

# `rw`: Robins-Wang variance estimation for multiple imputation

<!-- badges: start -->
[![R-CMD-check](https://github.com/yu000434/RW-package-test/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/yu000434/RW-package-test/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `rw` package computes [Robins-Wang variance
estimates](https://doi.org/10.1093/biomet/87.1.113) for analyses of multiply
imputed data. It supports parametric chained-equation imputation and predictive
mean matching with donor-source correction.

## Installation

The package requires the development version of `mice` that records fitted
imputation models through `tasks = "train"`.

```r
remotes::install_github("amices/mice@dev")
remotes::install_github("yu000434/RW-package-test")
```

## Parametric imputation

The first step is to impute your data using `mice`. For parametric imputation,
the package supports `method = "norm"` and `method = "logreg"`. When calling
`mice`, set `tasks = "train"` to save the fitted imputation models needed by
`with_rw()` and `pool_rw()`. The example below uses the `nhanes` data from
the `mice` package. After standardizing the variables, it creates five imputed
datasets using the `norm` method.

```{r parametric, message=FALSE}
library(rw)
library(mice)

nhanes_scaled <- as.data.frame(scale(nhanes))

set.seed(1)
imp_norm <- mice(nhanes_scaled, method = "norm", m = 5,
                 tasks = "train", print = FALSE)
```

Use `with_rw()` to fit a linear regression of `bmi` on `age` and `hyp` in each
completed dataset.

```{r parametric-fit}
fit_norm <- with_rw(imp_norm, lm(bmi ~ age + hyp))
```

Use `pool_rw()` to combine the coefficient estimates and compute Robins-Wang
standard errors and confidence intervals.

```{r parametric-pool}
pool_rw(fit_norm)
```

For comparison, Rubin's rules can be applied to the same imputations.

```{r rubin-parametric}
pool(with(imp_norm, lm(bmi ~ age + hyp))) |>
  summary()
```

## Predictive mean matching

For predictive mean matching, set the imputation method to `"pmmrw"` for
the incomplete variable and use `tasks = "train"` in `mice`. This method
uses the PMM donor draw implemented in `mice` and saves the donor IDs and
fitted matching models needed for variance
estimation. The correction accounts for repeated use of the same donor;
it does not change the imputed values. The example below imputes `bmi` using
`age` as a predictor, with the default pool of five donors.

```{r pmm, message=FALSE}
nhanes_pmm <- nhanes_scaled[c("age", "bmi")]

set.seed(2)
imp_pmm <- mice(nhanes_pmm, method = c(age = "", bmi = "pmmrw"),
                m = 5, tasks = "train", print = FALSE)
```

Next, use `with_rw()` to regress `bmi` on `age` in each completed dataset.
Because the PMM variable is the response of a linear model, `pool_rw()`
computes the PMM cross term and applies the donor-source correction
automatically.

```{r pmm-fit}
fit_pmm <- with_rw(imp_pmm, lm(bmi ~ age))
pool_rw(fit_pmm)
```

Using the same seed with `method = "pmm"` gives the same completed values.

```{r pmm-parity, message=FALSE}
set.seed(2)
imp_standard <- mice(nhanes_pmm, method = c(age = "", bmi = "pmm"),
                     m = 5, print = FALSE)
identical(imp_pmm$imp$bmi, imp_standard$imp$bmi)
```

The donor IDs used in each imputation are available when needed.

```{r donors}
head(extract_donor_id(imp_pmm, "bmi"))
```

## Giganti and Shepherd example

The package includes `giganti_data`, a simulated dataset based on
[Giganti and Shepherd (2020)](https://doi.org/10.1093/aje/kwaa153).
The variables `A` and `D` are observed only in a validation subsample of
1,000 individuals. The analysis model is a logistic regression of `D` on `A`,
with inclusion restricted to `A > 2`.

```{r gs-setup}
data("giganti_data", package = "rw")
gs <- giganti_data
predictors <- c("X1", "X2", "A.star", "D.star")

pred <- make.predictorMatrix(gs)
pred[,] <- 0L
pred["A", predictors] <- 1L
pred["D", c(predictors, "A")] <- 1L
```

With parametric imputation, `A` uses a normal model and `D` uses logistic
regression.

```{r gs-parametric, message=FALSE}
method <- make.method(gs)
method[] <- ""
method[c("A", "D")] <- c("norm", "logreg")

set.seed(41)
imp_gs_norm <- mice(gs, m = 2, method = method, predictorMatrix = pred,
                    tasks = "train", print = FALSE)
fit_gs_norm <- with_rw(
  imp_gs_norm,
  glm(D ~ A, subset = A > 2, family = binomial())
)
pool_rw(fit_gs_norm)
```

For predictive mean matching, `pmm_kappa_binomial()` computes the PMM cross
term for the binomial analysis before the final variance is assembled.

```{r gs-pmm, message=FALSE}
method[c("A", "D")] <- c("pmmrw", "logreg")

set.seed(42)
imp_gs_pmm <- mice(gs, m = 2, method = method, predictorMatrix = pred,
                   tasks = "train", print = FALSE)
fit_gs_pmm <- with_rw(
  imp_gs_pmm,
  glm(D ~ A, subset = A > 2, family = binomial())
)
kappa <- pmm_kappa_binomial(fit_gs_pmm, "A", threshold = 2)
pool_rw(fit_gs_pmm, pmm_kappa = kappa)
```

## Current scope

Parametric imputation supports `norm` and `logreg` with `lm()` or binomial
`glm()` analyses.

PMM supports one numeric imputed variable with fully observed matching
predictors. Use `pool_rw()` for an untransformed PMM response in `lm()`, or
`pmm_kappa_binomial()` for the binomial setting shown above. Other PMM
analyses require an expected-score function passed to `pmm_kappa()`.
See `?pool_rw` and `?pmm_kappa` for the supported model structures, and
`?mice.impute.pmmrw` for PMM options.
