---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
devtools::load_all(".", quiet = TRUE)
```

# `rw`: Robins-Wang variance estimation for multiple imputation

<!-- badges: start -->
[![R-CMD-check](https://github.com/yu000434/RW-package-test/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/yu000434/RW-package-test/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `rw` package computes [Robins-Wang variance
estimates](https://doi.org/10.1093/biomet/87.1.113) for analyses of multiply
imputed data. It supports parametric chained-equation imputation and predictive
mean matching with donor-source correction.

## Installation

The package requires the development version of `mice` that records fitted
imputation models through `tasks = "train"`.

```r
remotes::install_github("amices/mice@dev")
remotes::install_github("yu000434/RW-package-test")
```

## Parametric imputation

The first step is to impute your data using `mice`. For parametric imputation,
the package supports `method = "norm"` and `method = "logreg"`. When calling
`mice`, set `tasks = "train"` to save the fitted imputation models needed by
`with_rw()` and `pool_rw()`. The example below uses the `nhanes` data from
the `mice` package. After standardizing the variables, it creates five imputed
datasets using the `norm` method.

```{r parametric, message=FALSE}
library(rw)
library(mice)

nhanes_scaled <- as.data.frame(scale(nhanes))

set.seed(1)
imp_norm <- mice(nhanes_scaled, method = "norm", m = 5,
                 tasks = "train", print = FALSE)
```

Use `with_rw()` to fit a linear regression of `bmi` on `age` and `hyp` in each
completed dataset.

```{r parametric-fit}
fit_norm <- with_rw(imp_norm, lm(bmi ~ age + hyp))
```

Use `pool_rw()` to combine the coefficient estimates and compute Robins-Wang
standard errors and confidence intervals.

```{r parametric-pool}
pool_rw(fit_norm)
```

For comparison, Rubin's rules can be applied to the same imputations.

```{r rubin-parametric}
pool(with(imp_norm, lm(bmi ~ age + hyp))) |>
  summary()
```

## Predictive mean matching

For predictive mean matching, use `method = "pmmrw"` in `mice` and set
`tasks = "train"`. This method makes the same donor draws as MICE
PMM and saves the donor IDs and fitted matching models needed for variance
estimation. The correction accounts for repeated use of the same donor;
it does not change the imputed values. The example below imputes `bmi` using
`age` as a predictor, with the default pool of five donors.

```{r pmm, message=FALSE}
nhanes_pmm <- nhanes_scaled[c("age", "bmi")]
method <- make.method(nhanes_pmm)
method[] <- ""
method["bmi"] <- "pmmrw"

set.seed(2)
imp_pmm <- mice(nhanes_pmm, method = method, m = 5,
                tasks = "train", print = FALSE)
```

Next, use `with_rw()` to regress `bmi` on `age` in each completed dataset.
Because the PMM variable is the response of a linear model, `pool_rw()`
computes the PMM cross term and applies the donor-source correction
automatically.

```{r pmm-fit}
fit_pmm <- with_rw(imp_pmm, lm(bmi ~ age))
pool_rw(fit_pmm)
```

Using the same seed with MICE PMM gives the same completed values.

```{r pmm-parity, message=FALSE}
method["bmi"] <- "pmm"
set.seed(2)
imp_standard <- mice(nhanes_pmm, method = method, m = 5, print = FALSE)
identical(imp_pmm$imp$bmi, imp_standard$imp$bmi)
```

The donor IDs used in each imputation are available when needed.

```{r donors}
head(extract_donor_id(imp_pmm, "bmi"))
```

## Giganti and Shepherd example

The package includes `giganti_data`, a simulated dataset based on
[Giganti and Shepherd (2020)](https://doi.org/10.1093/aje/kwaa153).
The example below uses all 4,000 observations. Validated values of `A` and `D`
are available for 1,000 observations and are missing for the remainder.
The analysis fits a logistic regression of `D` on `A` among observations
with `A > 2`.

```{r gs-setup}
data("giganti_data", package = "rw")
gs <- giganti_data
predictors <- c("X1", "X2", "A.star", "D.star")

pred <- make.predictorMatrix(gs)
pred[,] <- 0L
pred["A", predictors] <- 1L
pred["D", c(predictors, "A")] <- 1L
```

With parametric imputation, `A` uses a normal model and `D` uses logistic
regression.

```{r gs-parametric, message=FALSE}
method <- make.method(gs)
method[] <- ""
method[c("A", "D")] <- c("norm", "logreg")

set.seed(41)
imp_gs_norm <- mice(gs, m = 2, method = method, predictorMatrix = pred,
                    tasks = "train", print = FALSE)
fit_gs_norm <- with_rw(
  imp_gs_norm,
  glm(D ~ A, subset = A > 2, family = binomial())
)
pool_rw(fit_gs_norm)
```

For predictive mean matching, `pmm_kappa_binomial()` computes the PMM cross
term for the binomial analysis before the final variance is assembled.

```{r gs-pmm, message=FALSE}
method[c("A", "D")] <- c("pmmrw", "logreg")

set.seed(42)
imp_gs_pmm <- mice(gs, m = 2, method = method, predictorMatrix = pred,
                   tasks = "train", print = FALSE)
fit_gs_pmm <- with_rw(
  imp_gs_pmm,
  glm(D ~ A, subset = A > 2, family = binomial())
)
kappa <- pmm_kappa_binomial(fit_gs_pmm, "A", threshold = 2)
pool_rw(fit_gs_pmm, pmm_kappa = kappa)
```

## Current scope

Parametric imputation supports `norm` and `logreg` with unweighted `lm()` or
binomial `glm()` analyses using the logit link.

PMM supports one numeric imputed variable with fully observed matching
predictors. Use `pool_rw()` for an untransformed PMM response in `lm()`, or
`pmm_kappa_binomial()` for the binomial setting shown above. Other PMM
analyses require an expected-score function passed to `pmm_kappa()`.
See `?pool_rw` and `?pmm_kappa` for the supported model structures, and
`?mice.impute.pmmrw` for PMM options.
